---
title: "p8105_hw5_jl7035"
author: "Jeffrey Lin"
date: "2024-11-11"
output: github_document
---

# Load Libraries
```{r}
library(tidyverse)

```


# Problem 1

## Functions for Randomly Drawing Birthdays and calculating prob of duplicates

```{r}
dup_bdays = function(samp_size) {
  bday_vec <- sample(c(1:365), size = samp_size, replace = TRUE)

  if (n_distinct(bday_vec) < samp_size) {
    return(TRUE)
  } else {
    return(FALSE)
  } 
}

```

## Run Simulations from sample size 2 to 50
```{r}

prob_by_size <- c()

for (sample_size in 2:50){
  num_dups = 0
  for (iteration in 1:10000) {
    if (dup_bdays(sample_size))
      num_dups = num_dups + 1
  }
  
  avg_prob = num_dups / 10000
  prob_by_size = append(prob_by_size, values = avg_prob)
}

dup_bday_tbl <-
  tibble(
    sample_size = 2:50,
    avg_prob = prob_by_size
  )

```


## Plot of probability of duplicate birthdays by sample size
```{r}
dup_bday_tbl %>% 
  ggplot(aes(x = sample_size, y = avg_prob)) +
  geom_point() +
  xlab("Sample Size") +
  ylab("Probability of Duplicate Birthdays") +
  ggtitle("Probability of Duplicate Birthdays by Sample Size")


```
As the sample size increases, the probability of duplicate birthdays increases.
The probability of having duplicate birthdays grows above 0.50 around a sample
size of 23. 

# Problem 2


## Generate 5000 datasets for mu = 0 and save p-values
```{r}
n <- 30
sigma <- 5
pop_param <- 0

x_bar_vec <- c()
p_val_vec <- c()

for (iter in 1:5000) {
  rand_samp <- rnorm(n, pop_param, sigma)
  x_bar <- mean(rand_samp)
  x_bar_vec = append(x_bar_vec, values = x_bar)
  
  p_val <- 
    t.test(rand_samp, mu = pop_param, conf.level = 0.95) %>%
    broom::tidy() %>% 
    pull(p.value)
    
  p_val_vec = append(p_val_vec, values = p_val)
  
}

t_test_results <- tibble(
  samp_avg = x_bar_vec,
  p_val = p_val_vec
)


```


# Write Function that handles generating datasets and finding p-values for mu
```{r}
gen_test_data <- function(pop_param_vec, n, sigma ) {
  x_bar_vec <- c()
  p_val_vec <- c()
  for (val in pop_param_vec) {
    for (iter in 1:5000) {
      rand_samp <- rnorm(n, val, sigma)
      x_bar <- mean(rand_samp)
      x_bar_vec <- c(x_bar_vec, x_bar)

      p_val <- 
      t.test(rand_samp, mu = val, conf.level = 0.95) %>%
      broom::tidy() %>% 
      pull(p.value)
      p_val_vec <- c(p_val_vec, p_val)
      
  t_test_results <- bind_rows(t_test_results, tibble(x_bar_vec, p_val_vec))
    }
  }
}

```

## Generate datasets and perform t-tests for mu values 1 to 5
```{r}
mu_vec <- c(1, 2, 3, 4, 5)
gen_test_data(mu_vec, n, sigma)

```

